---
title: "SPARQL with R"
author: "Katharina Kaelin"
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: true
    theme: lumen
    highlight: tango
---

<style type="text/css">
.main-container {
  max-width: 1800px;
  margin-left: 0;
  margin-right: auto;
}
blockquote {
    padding: 10px 20px;
    margin: 0 0 20px;
    font-size: 14px;
    border-left: 5px solid #eee;
}

h1.title {
  font-size: 38px;
  color: #000000;
  }
  
h1 { /* Header 1 */
  font-size: 28px;
  color: #0033cc;
}
h2 { /* Header 2 */
    font-size: 28px;
	color: #0099ff
}
h3 { /* Header 3 */
    font-size: 14px;
	color: #6600cc
}
</style>

https://github.com/statistikZH/R_AS_GIS <br>
https://query.wikidata.org/ <br>
rm(list = ls())

# Introduction

**An overview of R libraries to query Wikidata (27.1.2019)** <br>
https://www.lehir.net/how-to-query-wikidata-in-r/ <br> 
=> https://www.lehir.net/how-to-query-wikidata-in-r/#summary


**> WikidataR** <br>
https://github.com/Ironholds/WikidataR <br>
https://cran.r-project.org/web/packages/WikidataR/index.html <br>
https://cran.r-project.org/web/packages/WikidataR/WikidataR.pdf 

**> WikidataQueryServiceR** <br>
https://github.com/bearloga/WikidataQueryServiceR <br>
https://cran.r-project.org/web/packages/WikidataQueryServiceR/index.html <br>
https://cran.r-project.org/web/packages/WikidataQueryServiceR/WikidataQueryServiceR.pdf 

**> SPARQL** <br>
https://cran.r-project.org/web/packages/SPARQL/index.html <br>
https://cran.r-project.org/web/packages/SPARQL/SPARQL.pdf 

# R Setup

```{r setup, echo=T, results='hide', message=FALSE, warning=FALSE}

# Import libraries
library(WikidataQueryServiceR)
library(SPARQL)
library(tidyverse)
library(sf)
library(stringr)

# Number formatting
options(scipen = 1000000)
options(digits = 6)

```

# Museum in Canton Zürich

https://w.wiki/AwP

```{r, collapse=TRUE, warning=FALSE,  message=FALSE}
## WikidataQueryServiceR
start.time <- Sys.time()
museum1_df <- query_wikidata('SELECT DISTINCT ?item ?name ?coord ?lat ?lon
WHERE
{
hint:Query hint:optimizer "None" .
?item wdt:P131* wd:Q11943 .
?item wdt:P31/wdt:P279* wd:Q33506 .
?item wdt:P625 ?coord .
?item p:P625 ?coordinate .
?coordinate psv:P625 ?coordinate_node .
?coordinate_node wikibase:geoLatitude ?lat .
?coordinate_node wikibase:geoLongitude ?lon .
SERVICE wikibase:label {
bd:serviceParam wikibase:language "ca" .
?item rdfs:label ?name
}
}
ORDER BY ASC (?name)')
end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken

## SPARQL 
start.time <- Sys.time()
endpoint <- "https://query.wikidata.org/sparql"
query <- 'SELECT DISTINCT ?item ?name ?coord ?lat ?lon
WHERE
{
hint:Query hint:optimizer "None" .
?item wdt:P131* wd:Q11943 .
?item wdt:P31/wdt:P279* wd:Q33506 .
?item wdt:P625 ?coord .
?item p:P625 ?coordinate .
?coordinate psv:P625 ?coordinate_node .
?coordinate_node wikibase:geoLatitude ?lat .
?coordinate_node wikibase:geoLongitude ?lon .
SERVICE wikibase:label {
bd:serviceParam wikibase:language "ca" .
?item rdfs:label ?name
}
}
ORDER BY ASC (?name) '
museum2 <- SPARQL(endpoint,query,curl_args=list(useragent=R.version.string))
museum2_df <- museum2$results
end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken
```

# Municipalities of Canton Zürich

http://yasgui.org/short/fg7fNak6G

```{r, collapse=TRUE, warning=FALSE,  message=FALSE}
## SPARQL 
endpoint <- "https://ld.geo.admin.ch/query"
query <- 'PREFIX schema: <http://schema.org/>
PREFIX gn: <http://www.geonames.org/ontology#>
PREFIX wdt: <http://www.wikidata.org/prop/direct/>
PREFIX wd: <http://www.wikidata.org/entity/>
PREFIX wikibase: <http://wikiba.se/ontology#>
PREFIX geo: <http://www.opengis.net/ont/geosparql#>
PREFIX dct: <http://purl.org/dc/terms/>
PREFIX st: <https://ld.geo.admin.ch/def/>
select ?Municipality ?Name ?Population ?bfs ?WKT
where{
?Municipality gn:featureCode gn:A.ADM3 .
?Municipality schema:name ?Name .
?Municipality gn:population?Population .
?Municipality st:bfsNumber ?bfs .
?Municipality dct:issued ?Date .
?Municipality gn:parentADM1 ?InCanton .
?InCanton schema:name ?CantonName .
?Municipality geo:hasGeometry ?Geometry .
?Geometry geo:asWKT ?WKT .
FILTER (?Date = "2019-01-01"^^xsd:date)
FILTER (?CantonName = "Zürich")
}'
municipality <- SPARQL(endpoint,query)
municipality_df <- municipality$results
```


# Convert df to spatial df

```{r, collapse=TRUE, warning=FALSE,  message=FALSE}
## museum
museum_sf <- sf::st_as_sf(x = museum2_df, coords = c("lon", "lat"), crs= 4326) # epsg:4326 is the ID of wgs84

## municipality
municipality_df$WKT[1]
municipality_df$WKT_corrected <-str_match(municipality_df$WKT, '\"(.*?)\"')[,2]
municipality_df$WKT_corrected[1]
municipality_sfc = st_as_sfc(municipality_df$WKT_corrected, crs=4326) # epsg:4326 is the ID of wgs84
municipality_sf = st_sf(municipality_df %>% select(-WKT, -WKT_corrected), geometry =municipality_sfc) 

# Plot result: R base plot
plot(st_geometry(municipality_sf))
plot(st_geometry(museum_sf), pch = 19, col="blue", cex = 0.5, add = TRUE)
plot(st_geometry(municipality_sf), add = TRUE)
legend(x=2750000,y=1310000,
       c("Museum","Muncipality"),
       lty=c(NA,1),
       pch=c(19,NA),
       cex=.8, 
       col=c("blue","black"),
       bty='n'
       )
```

# Calculate density of museums per municipality

```{r, collapse=TRUE, warning=FALSE,  message=FALSE}
# Spatial Join: instead of joining dataframes via an equal ID we join data- frames based on an equal location. 
spjoin_sf <- sf::st_join(museum_sf, municipality_sf)
spjoin_sf

# Density calculation

# > 1. Count points per polygon
pts_count <- spjoin_sf %>%
   dplyr::group_by(bfs) %>%
   dplyr::summarise(count=n())
municipality_sf <- municipality_sf %>%
  dplyr::left_join(pts_count %>% st_set_geometry(NULL) , by = c("bfs" ))

# > 2. Calculate area of polygon
municipality_sf <- municipality_sf %>%
  dplyr::mutate(mun_area_m2 =as.vector(sf::st_area(.)))

# > 3. Calculate density: count/area
municipality_sf$density <- municipality_sf$count / municipality_sf$mun_area_m2 * 1000000

# Plot result: tmap
# > tmap static
tmap::tmap_mode
tmap::tm_shape(municipality_sf) +
  tmap::tm_fill("density",
          title="Density of public transportation stops",
          style="quantile", 
          palette="BuGn", 
          colorNA = "grey") +
  tmap::tm_borders() +
  tmap::tm_layout(frame = FALSE,
                  legend.position = c("right", "top"),
                  legend.outside = TRUE)
# Export Data as shp
st_write(museum_sf, "./museum.shp", delete_layer = TRUE)
st_write(municipality_sf, "./municipality.shp", delete_layer = TRUE)

```

